---
title: "assignmet 12"
author: "Alexander Dyakin"
date: "2023-12-04"
output: pdf_document
---
In this I will take a image from my high school graduation and use svd compression to reduce the size of the image and compare the diffrence in quality

need these libraries to view the image
```{r}
if(!require("pixmap"))
    install.packages("pixmap")

if(!require("bmp"))
    install.packages("bmp")

```
```{r}
coolImage = read.bmp("coolAlex.bmp")
dim(coolImage)
## now use pixmap to graph the image:
coolImage = pixmapGrey(coolImage)
print("this may take a minute...")
plot(coolImage)
```
that took WAAAAYYYY to long
lets find the eigen values for the matrix of grey pixels
```{r}
#let M be the matrix of all grey values of each pixel
M = coolImage@grey
## first find M transpose M
MTM = t(M) %*% M
## use eigen to find eigen values of the matrx
eigen = eigen(MTM)
```
check how many eigen values we found and order them from largest to smallest
```{r}
print("number of eigen values:")
print(length(eigen$values))
orderedEigen = order(eigen$values,decreasing = TRUE)
orderedEigen = eigen$values[orderedEigen]
```
 a lot of eigen values are close to zero so lets save time and recorces and just make them zero
```{r}
orderedEigen[which(orderedEigen < 1e-13)] =0
```
make sure all eigen values are positive. then see how many eigen values there are
```{r}
index = which(eigen$values>1e-13)
posEigenOrdered = order(eigen$values[index],decreasing = TRUE)
index = index[posEigenOrdered]
print("number of eigen values:")
print(length(posEigenOrdered))
```
now to use the eigen values to find the sigular values by taking their sqrt
```{r}
singularValues = sqrt(eigen$values[posEigenOrdered])
print("number of singular values")
length(singularValues)
```
set diagonal vector D to be the diagonal of the singular values. set V to be the vector of the normalized eigen vectors
```{r}
D = diag(singularValues)
V = eigen$vectors[index,index]
```
compute the SVD
```{r}
c1 = colSums(V)
c2 = matrix(rep(1,dim(V)[1]),nr=1)
sum(abs(c1-c2))

for (i in 1:dim(V)[1]) {
    V[,i] = V[,i]/c1[i]
}
##colSums(V)
```
function that will allow us to use the SVD to change the quality of the image
```{r}
compress_image = function(image, n) {
    # Given an image svd 
    # use only the first n singular values
  out = list() ## return type of list
  ##make sure the dimensions match for the number of singular values possible
    if (n > length(image$d)) {
      ## n is too large, set it to the max possible
        n = length(image$d)
    }
  ## get matrcies D V and V from native svd result
    d_tmp = image$d[1:n]
    u_tmp = image$u[,1:n]
    v_tmp = image$v[,1:n]
    # First, compute the resulting image
    out$img = mat.or.vec(nr = dim(image$u)[1], nc = dim(image$v)[1])
    for (i in 1:n) {
        out$img = out$img + d_tmp[i] * u_tmp[,i] %*% t(v_tmp[,i]) 
    }
    # remove any values that are less than 0 because we cant have negitive color since this is just black and white
    if (min(min(out$img)) < 0 ) {
      ## replace those values with the "opposite" of that value
        out$img = out$img - min(min(out$img))
    }
    out$img = out$img / max(max(out$img))
    # Store some information: number of points needed and percentage of the original required
    # number of original pixels in the image
    out$numOriginal = dim(image$u)[1] * dim(image$v)[2]
    # number of pixels used after compression
    out$numCompressed = length(d_tmp) + dim(u_tmp)[1]*dim(u_tmp)[2] + dim(v_tmp)[1]*dim(v_tmp)[2] 
    out$pct_of_original = out$numCompressed / out$numOriginal * 100
    # Return the result
    return(out)
}
```
use native SVD function to find the svd. then find the image so its less detailed and takes less space
```{r}
print("this may take a minute...")
my_SVD = svd(M)

```
 compress the image and see how much space we saved
```{r}
newImage = coolImage
compressed = compress_image(my_SVD,30)
newImage@grey = compressed$img
plot(newImage)
print("number of pixels in the original image:")
compressed$numOriginal
print("number of pixels used in the compressed image:")
compressed$numCompressed
```
now lets try with a really small number of singular values
```{r}
newImage = coolImage
compressed = compress_image(my_SVD,9)
newImage@grey = compressed$img
plot(newImage)
print("number of pixels in the original image:")
compressed$numOriginal
print("number of pixels used in the compressed image:")
compressed$numCompressed
```
look at that hansom little potato man
P.S my senior quote under that image was:
"dont you dare use that as your senior photo" -mom
i was on the front page of reddit for that :)
